% !TEX TS-program = pdflatexmk
\documentclass[12pt]{article}

% Layout.
\usepackage[top=.75in, bottom=0.75in, left=.75in, right=.75in, headheight=1in, headsep=6pt]{geometry}

% Fonts.
\usepackage{mathptmx}
\usepackage[scaled=0.86]{helvet}
\renewcommand{\emph}[1]{\textsf{\textbf{#1}}}

% Misc packages.
\usepackage{amsmath,amssymb,latexsym}
\usepackage{graphicx,tikz}
\usepackage{array}
\usepackage{xcolor}
\usepackage{multicol}
\usepackage{tabularx,colortbl}
\usepackage{enumitem}
%to make tikz pics work
\usepackage{tikz,pgfplots}
\usetikzlibrary{arrows}
\newcommand{\midarrow}{\tikz \draw[-triangle 90] (0,0) -- +(.1,0);}

\usepackage[colorlinks=true]{hyperref}

% Paragraph spacing
\parindent 0pt
\parskip 6pt plus 1pt
\def\tableindent{\hskip 0.5 in}
\def\ts{\hskip 1.5 em}

\usepackage{fancyhdr}
\pagestyle{fancy} 
\lhead{\large\sf\textbf{MATH 405}}
\rhead{\large\sf\textbf{Spring 2024}}
\chead{\large\sf\textbf{HW 10}}

\newcommand{\localhead}[1]{\par\smallskip\textbf{#1}\nobreak\\}%
\def\heading#1{\localhead{\large\emph{#1}}}
\def\subheading#1{\localhead{\emph{#1}}}

%% Special Math Symbol shortcuts
\newcommand{\bbN}{\mathbb{N}}
\newcommand{\bbZ}{\mathbb{Z}}
\newcommand{\bbR}{\mathbb{R}}
\newcommand{\bbQ}{\mathbb{Q}}
\newcommand{\bbC}{\mathbb{C}}
\newcommand{\bbM}{\mathbb{M}}


\newcommand{\rad}{\text{rad}}
\newcommand{\diam}{\text{diam}}

\usetikzlibrary{calc,arrows.meta}
\usetikzlibrary{arrows}
\newcommand{\marrow}{\tikz \draw[-triangle 90] (0,0) -- +(.1,0);}


\begin{document}
\begin{enumerate}
\item Find all of  the ideals in each of the following rings. Which of these ideals are maximal and which are prime?
	\begin{enumerate}
	\item $\bbZ_{18}$
	\item $\bbZ_{25}$
	\item $\mathbb{M}_2(\bbR)$
	\item $\bbQ$
	\end{enumerate}
\item Find all homomorphisms $\phi: \mathbb{Z}/6\mathbb{Z} \to  \mathbb{Z}/15\mathbb{Z}.$
\item Prove that $\bbR$ is not isomorphic to $\bbC.$
\item Define a map $\phi: \bbC \to \bbM_2(\bbR)$ by $\phi(a+bi)=\begin{pmatrix} a&b\\-b&a \end{pmatrix}$. Show that $\phi$ is an isomorphism of $\bbC$ to its image under $\phi.$
\item Prove that the Gaussian integers, $\bbZ[i],$ form an integral domain.
\item Prove that if $R$ is a field, the only ideals of $R$ are $\{0\}$ and $R$ itself.
\item Suppose $\phi: R \to S$ is a ring homomorphism. Prove each of the following statements.
	\begin{enumerate}
	\item If $R$ is a commutative ring, then $\phi(R)$ is a commutative ring.
	\item $\phi(0_R)=0_S.$
	\item If $\phi$ is onto, then $\phi(1_R)=1_S.$
	\item If $R$ is a field and $\phi(R)\not=\{0\},$ then $\phi(R)$ is a field.
	\end{enumerate}
\item Let $R$ be a ring and let $\{R_\alpha\}$ be a collection of subrings of $R$ for $\alpha \in A.$ Prove that $ \displaystyle\cap_{\alpha \in A} R_\alpha$ is also a subring and show by example that the union of two subrings may not be a subring. Note that $A$ is just an index set. 
\item Let $R$ be a ring and let $\{I_\alpha\}$ be a collection of ideals of $R$ for $\alpha \in A.$ Prove that $ \displaystyle\cap_{\alpha \in A} I_\alpha$ is also an ideal and show by example that the union of two ideals may not be an ideal. Note that $A$ is just an index set.
\end{enumerate}
\end{document}

	
	
	
	
	
	

 

