% !TEX TS-program = pdflatexmk
\documentclass[12pt]{article}

% Layout.
\usepackage[top=.75in, bottom=0.75in, left=.75in, right=.75in, headheight=1in, headsep=6pt]{geometry}

% Fonts.
\usepackage{mathptmx}
%\usepackage[scaled=0.86]{helvet}
\renewcommand{\emph}[1]{\textsf{\textbf{#1}}}
\renewcommand{\familydefault}{\sfdefault}
%Syl adendum
\usepackage[colorlinks = true,linkcolor = blue, urlcolor  = blue]{hyperref}
\def\mailto#1{\href{mailto:#1}{#1}}

% Misc packages.
\usepackage{amsmath,amssymb,latexsym,multicol}
\usepackage{graphicx}
\usepackage{array}
\usepackage{xcolor}
\usepackage{multicol}
\usepackage{tabularx,colortbl}
\usepackage{enumitem}
%to make tikz pics work
\usepackage{tikz,pgfplots}

\usepackage[colorlinks=true]{hyperref}

%% Special Math Symbol shortcuts
\newcommand{\bbN}{\mathbb{N}}
\newcommand{\bbZ}{\mathbb{Z}}
\newcommand{\bbR}{\mathbb{R}}
\newcommand{\bbQ}{\mathbb{Q}}

% Paragraph spacing
\parindent 0pt
\parskip 6pt plus 1pt
\def\tableindent{\hskip 0.5 in}
\def\ts{\hskip 1.5 em}

\usepackage{fancyhdr}
\pagestyle{fancy} 
\lhead{\large\sf\textbf{MATH 405 }}
\rhead{\large\sf\textbf{Spring 2024}}
\chead{\large\sf\textbf{Midterm III}}

\newcommand{\localhead}[1]{\par\smallskip\textbf{#1}\nobreak\\}%
\def\heading#1{\localhead{\large\emph{#1}}}
\def\subheading#1{\localhead{\emph{#1}}}

\newenvironment{clist}%
{\bgroup\parskip 0pt\begin{list}{$\bullet$}{\partopsep 4pt\topsep 0pt\itemsep -2pt}}%
{\end{list}\egroup}%

\usetikzlibrary{calc}
\pgfplotsset{my style/.append style={axis x line=middle, axis y line=
middle, xlabel={$x$}, ylabel={$y$}, axis equal }}


\begin{document}
\quad
\vskip 2cm
\strut\vtop{\halign{\emph#\hskip 0.5em\hfil&#\hbox to 2in{\hrulefill}\cr
\emph{\fontsize{18}{22}\selectfont Name:}&\cr
\noalign{\vskip 10pt}}}

\vfill
{\fontsize{18}{22}\selectfont\emph{Rules:}}

You have 1.5 hours to complete this midterm.  

Partial credit will be awarded, but you must show your work.

No notes, books, or other aids are allowed.

Turn off anything that might go beep during the exam.

Good luck!
\vfill
\def\emptybox{\hbox to 2em{\vrule height 16pt depth 8pt width 0pt\hfil}}
\def\tline{\noalign{\hrule}}
\centerline{\vbox{\offinterlineskip
{
\bf\sf\fontsize{18pt}{22pt}\selectfont
\hrule
\halign{
\vrule#&\strut\quad\hfil#\hfil\quad&\vrule#&\quad\hfil#\hfil\quad
&\vrule#&\quad\hfil#\hfil\quad&\vrule#\cr
height 3pt&\omit&&\omit&&\omit&\cr
&Problem&&Possible&&Score&\cr\tline
height 3pt&\omit&&\omit&&\omit&\cr
&1&&16&&\emptybox&\cr\tline
&2&&20&&\emptybox&\cr\tline
&3&&16&&\emptybox&\cr\tline
&4&&12&&\emptybox&\cr\tline
&5&&12&&\emptybox&\cr\tline
&6&&24&&\emptybox&\cr\tline
&Extra Credit&&5&&\emptybox&\cr\tline
&Total&&100&&\emptybox&\cr
}\hrule}}}

\newpage
\begin{enumerate}
\item Let $G$ and $H$ be groups and let $\phi: G \to H$ be a group homomorphism.
	\begin{enumerate}
	\item (2 pts) State the definition of a \emph{group homomorphism}.
	\vfill
	\item (2 pts) State the definition of the \emph{kernel of $\phi,$} $\text{ker } \phi.$ 
	\vfill
	\item (12 pts) Prove $\text{ker }\phi$ is a normal subgroup of $G.$ (Note that you must show $\text{ker }\phi$ is a subgroup of $G$ \emph{and} that it is normal.)
	\vspace{6.5in}
	\end{enumerate}
\newpage
%examples
\item (20 points) Give an examples of the following, if they exist. Otherwise briefly explain why such examples do not exist.
	\begin{enumerate}
	\item A commutative ring with unity that is not an integral domain.
	\vfill
	\item A ring that is an integral domain but is not a field.
	\vfill
	\item A ring $R$ and a nontrivial subring $I$ such that $I$ is an ideal of $R$
	\vfill
	\item A ring $R$ and a nontrivial subring $S$ such that $S$ is \emph{not} an ideal of $R$
	\vfill
	\item A ring $R$ and an ideal $I$ that is prime.
	\vfill
	\end{enumerate}
\newpage

%%hwk # 9 + First Isom Theorem
\item (16 points) 
	\begin{enumerate}
	\item (4 pts) State the First Isomorphism Theorem (for groups)
	\vspace{2in}
	\item (12 pts) Let $\psi: G \to H$ be a group homomorphism. Prove that $\psi$ is one-to-one if and only if $\psi^{-1}(e_H)=\{e_G\}.$\\	
	\end{enumerate}
\newpage

\item (12 points) Prove that if $R$ is a field, the only ideals of $R$ are $\{0\}$ and $R$ itself.\\
\vfill

%
\item (12 points) Let $R$ be a ring and let $a \in R.$ Prove that the set $S=\{r \in R \, : \, ra=0\}$ is a subring of $R.$ Note that you should not assume $R$ is commutative.\\

\vfill
\newpage
%short answer
\item (24 points) 
\begin{enumerate}
	\item List all nonisomorphic abelian groups of order 24.
	\vfill
	\item Let $\mathbb{R}$ be the ring of real numbers under the usual operations of addition and multiplication. Explain why the function $f:\mathbb{R} \to \mathbb{R}$ defined as $f(x)=2x+1$ is not ring homomorphism.
	\vfill
	\item Find all group homomorphisms from $\mathbb{Z}_{16}$ to $\mathbb{Z}_{18}.$ Your answer(s) must be stated as functions.
	\vfill
	\item Give a maximal ideal in the ring $\mathbb{Z}_{20}$
	\vfill
\end{enumerate}
\end{enumerate}
\newpage
\emph{Extra Credit:} (5 points) Prove that every finite integral domain is a field.
\vfill

\end{document}

Let $\mathbb{M}_2(\mathbb{R})$ be the ring of $2 \times 2$ matrices under matrix addition and matrix multiplication.
	\begin{enumerate}
	\item What is the additive identity of $\mathbb{M}_2(\mathbb{R})$?
	\vfill
	\item Does $\mathbb{M}_2(\mathbb{R})$ have a multiplicative identity? Explain.
	\vfill
	\item Does $\mathbb{M}_2(\mathbb{R})$ have any zero divisors?  Explain.
	\vfill
	\item What are the units of $\mathbb{M}_2(\mathbb{R})$? Describe them or explain why none exist?
	\vfill
	\end{enumerate}