% !TEX TS-program = pdflatexmk
\documentclass[12pt]{article}

% Layout.
\usepackage[top=.75in, bottom=0.75in, left=.75in, right=.75in, headheight=1in, headsep=6pt]{geometry}

% Fonts.
\usepackage{mathptmx}
%\usepackage[scaled=0.86]{helvet}
\renewcommand{\emph}[1]{\textsf{\textbf{#1}}}
\renewcommand{\familydefault}{\sfdefault}
%Syl adendum
\usepackage[colorlinks = true,linkcolor = blue, urlcolor  = blue]{hyperref}
\def\mailto#1{\href{mailto:#1}{#1}}

% Misc packages.
\usepackage{amsmath,amssymb,latexsym,multicol}
\usepackage{graphicx}
\usepackage{array}
\usepackage{xcolor}
\usepackage{multicol}
\usepackage{tabularx,colortbl}
\usepackage{enumitem}
%to make tikz pics work
\usepackage{tikz,pgfplots}

\usepackage[colorlinks=true]{hyperref}

%% Special Math Symbol shortcuts
\newcommand{\bbN}{\mathbb{N}}
\newcommand{\bbZ}{\mathbb{Z}}
\newcommand{\bbR}{\mathbb{R}}
\newcommand{\bbQ}{\mathbb{Q}}
\newcommand{\bbM}{\mathbb{M}}


% Paragraph spacing
\parindent 0pt
\parskip 6pt plus 1pt
\def\tableindent{\hskip 0.5 in}
\def\ts{\hskip 1.5 em}

\usepackage{fancyhdr}
\pagestyle{fancy} 
\lhead{\large\sf\textbf{MATH 405 }}
\rhead{\large\sf\textbf{Spring 2024}}
\chead{\large\sf\textbf{Final Exam}}

\newcommand{\localhead}[1]{\par\smallskip\textbf{#1}\nobreak\\}%
\def\heading#1{\localhead{\large\emph{#1}}}
\def\subheading#1{\localhead{\emph{#1}}}

\newenvironment{clist}%
{\bgroup\parskip 0pt\begin{list}{$\bullet$}{\partopsep 4pt\topsep 0pt\itemsep -2pt}}%
{\end{list}\egroup}%

\usetikzlibrary{calc}
\pgfplotsset{my style/.append style={axis x line=middle, axis y line=
middle, xlabel={$x$}, ylabel={$y$}, axis equal }}


\begin{document}
\quad
\vskip 2cm
\strut\vtop{\halign{\emph#\hskip 0.5em\hfil&#\hbox to 2in{\hrulefill}\cr
\emph{\fontsize{18}{22}\selectfont Name:}&\cr
\noalign{\vskip 10pt}}}

\vfill
{\fontsize{18}{22}\selectfont\emph{Rules:}}

You have 2 hours to complete this Final Exam.  

Partial credit will be awarded, but you must show your work.

No notes, books, or other aids are allowed.

Turn off anything that might go beep during the exam.

Good luck!
\vfill
\def\emptybox{\hbox to 2em{\vrule height 16pt depth 8pt width 0pt\hfil}}
\def\tline{\noalign{\hrule}}
\centerline{\vbox{\offinterlineskip
{
\bf\sf\fontsize{18pt}{22pt}\selectfont
\hrule
\halign{
\vrule#&\strut\quad\hfil#\hfil\quad&\vrule#&\quad\hfil#\hfil\quad
&\vrule#&\quad\hfil#\hfil\quad&\vrule#\cr
height 3pt&\omit&&\omit&&\omit&\cr
&Problem&&Possible&&Score&\cr\tline
height 3pt&\omit&&\omit&&\omit&\cr
&1&&10&&\emptybox&\cr\tline
&2&&20&&\emptybox&\cr\tline
&3&&16&&\emptybox&\cr\tline
&4&&12&&\emptybox&\cr\tline
&5&&12&&\emptybox&\cr\tline
&6&&24&&\emptybox&\cr\tline
&Extra Credit&&5&&\emptybox&\cr\tline
&Total&&100&&\emptybox&\cr
}\hrule}}}

\newpage
\begin{enumerate}
\item Let $G$ and $H$ be groups and let $\phi: G \to H$ be a group homomorphism.
	\begin{enumerate}
	\item ({\small 2 pts}) State the definition of a \emph{group homomorphism}.
	\vfill
	\item ({\small 2 pts}) State the definition of the \emph{kernel of $\phi,$} $\text{ker } \phi.$ 
	\vfill
	\item ({\small 8 pts}) Prove $\text{ker }\phi$ is a normal subgroup of $G.$ (Note that you must show $\text{ker }\phi$ is a subgroup of $G$ \emph{and} that it is normal.)
	\vspace{6.5in}
	\end{enumerate}
\newpage
%examples
\item ({\small18 points}) Give an examples of the following, if they exist. Otherwise, briefly explain why such examples do not exist.
	\begin{enumerate}
	\item An infinite nonabelian group.\\
	\vfill
	\item A nonabelian group of order $n=11.$\\
	\vfill
	\item An infinite group $G$ with multiple elements of finite order.\\  
	\vfill
	\item Three nonisomorphic groups of order 12.\\
	\vfill
	\item A commutative ring with unity that is not an integral domain.\\
	\vfill
	\item A ring $R$ and an ideal $I$ such that $I$ is prime.\\
	\vfill
	\item A ring $R$ and an ideal $I$ that is maximal in $R$.\\
	\vfill
	\item A ring $R$ such that $R[x]$ contains a unit of degree at least 1.\\
	\vfill
	\end{enumerate}
\newpage
\item ({\small 12 points}) Let $G$ be an abelian group. Let $H=\{ a \in G\, : \, |a| < \infty\}.$ (That is, $H$ consists of all elements of $G$ of finite order.)\\

Prove that $H$ is a subgroup of $G.$\\
\newpage
\item 
	\begin{enumerate}
	\item ({\small 4 points}) State Lagrange's Theorem
	\vspace{2in}
	\item  ({\small 10 points}) Let $G$ be a group of order $pq$ where $p$ and $q$ are both primes. Prove that every proper subgroup of $G$ is cyclic.
	\vfill
	\end{enumerate}
\newpage
\item Let $R$ be the ring of functions $f: \bbR \to \bbR$ with the usual operations of addition and multiplication. Let $S$ be the set of differentiable functions in $R.$ (Note: All the functions in $R$, and therefore $S$, have domain $\bbR.$)
	\begin{enumerate}
	\item ({\small 10 points}) Prove that $S$ is a subring of $R.$
	\vfill
	\item ({\small 4 points}) Prove that $S$ is \emph{not} an ideal of $R.$
	\vfill
	\end{enumerate}
\newpage
%\quad \\
\item Let $R=\left\{\begin{bmatrix} a&b\\0&c \end{bmatrix} \, : \, a,b,c \in \bbZ \right\}.$ Consider the function $\phi: R \to \bbZ$ defined by $\phi \left( \begin{bmatrix} a&b\\0&c \end{bmatrix} \right)=a.$
	\begin{enumerate}
	\item ({\small 10 points}) Prove that $\phi$ is a ring homomorphism.
	\vfill
	\item ({\small 4 points}) Determine the kernel of $\phi.$
	\vfill
	\end{enumerate}
\newpage
\item ({\small{ 4 points each}}) Short Answer
	\begin{enumerate}
	\item What is the order of the factor group $\bbZ_{60}/\langle 15 \rangle$?
	\vfill
	\item What is the order of the element $10+\langle 15 \rangle$ in the factor group $\bbZ_{60}/\langle 15 \rangle$?
	\vfill
	\item Is $2x^4+1$ an element of  $\langle x^2+2 \rangle,$  the ideal generated by $x^2+2$ in $\bbZ_3[x]$? Justify your answer.
	\vfill
	\item Show that the map $f(x)=5x$ is \emph{not} a ring homomorphism from $\bbZ_{12}$ to $\bbZ_{60}.$
	\vfill
	\end{enumerate}
\end{enumerate}

\textbf{5 pts Extra Credit:} Suppose $f(x)$ is irreducible in $F[x]$, where $F$ is a field. Prove that for every nonzero polynomial $g(x) \in F[x],$ either $\text{gcd}(f(x),g(x))=1$ or $f(x) \, \big\vert g(x).$\\

\vspace{3in}
\end{document}