% !TEX TS-program = pdflatexmk
\documentclass[12pt]{article}

% Layout.
\usepackage[top=.75in, bottom=0.75in, left=.75in, right=.75in, headheight=1in, headsep=6pt]{geometry}

% Fonts.
\usepackage{mathptmx}
%\usepackage[scaled=0.86]{helvet}
\renewcommand{\emph}[1]{\textsf{\textbf{#1}}}
\renewcommand{\familydefault}{\sfdefault}
%Syl adendum
\usepackage[colorlinks = true,linkcolor = blue, urlcolor  = blue]{hyperref}
\def\mailto#1{\href{mailto:#1}{#1}}

% Misc packages.
\usepackage{amsmath,amssymb,latexsym,multicol}
\usepackage{graphicx}
\usepackage{array}
\usepackage{xcolor}
\usepackage{multicol}
\usepackage{tabularx,colortbl}
\usepackage{enumitem}
%to make tikz pics work
\usepackage{tikz,pgfplots}

\usepackage[colorlinks=true]{hyperref}

%% Special Math Symbol shortcuts
\newcommand{\bbN}{\mathbb{N}}
\newcommand{\bbZ}{\mathbb{Z}}
\newcommand{\bbR}{\mathbb{R}}
\newcommand{\bbQ}{\mathbb{Q}}

% Paragraph spacing
\parindent 0pt
\parskip 6pt plus 1pt
\def\tableindent{\hskip 0.5 in}
\def\ts{\hskip 1.5 em}

\usepackage{fancyhdr}
\pagestyle{fancy} 
\lhead{\large\sf\textbf{MATH 405 }}
\rhead{\large\sf\textbf{Spring 2024}}
\chead{\large\sf\textbf{Midterm II}}

\newcommand{\localhead}[1]{\par\smallskip\textbf{#1}\nobreak\\}%
\def\heading#1{\localhead{\large\emph{#1}}}
\def\subheading#1{\localhead{\emph{#1}}}

\newenvironment{clist}%
{\bgroup\parskip 0pt\begin{list}{$\bullet$}{\partopsep 4pt\topsep 0pt\itemsep -2pt}}%
{\end{list}\egroup}%

\usetikzlibrary{calc}
\pgfplotsset{my style/.append style={axis x line=middle, axis y line=
middle, xlabel={$x$}, ylabel={$y$}, axis equal }}


\begin{document}
\quad
\vskip 2cm
\strut\vtop{\halign{\emph#\hskip 0.5em\hfil&#\hbox to 2in{\hrulefill}\cr
\emph{\fontsize{18}{22}\selectfont Name:}&\cr
\noalign{\vskip 10pt}}}

\vfill
{\fontsize{18}{22}\selectfont\emph{Rules:}}

You have 1.5 hours to complete this midterm.  

Partial credit will be awarded, but you must show your work.

No notes, books, or other aids are allowed.

Turn off anything that might go beep during the exam.

Good luck!
\vfill
\def\emptybox{\hbox to 2em{\vrule height 16pt depth 8pt width 0pt\hfil}}
\def\tline{\noalign{\hrule}}
\centerline{\vbox{\offinterlineskip
{
\bf\sf\fontsize{18pt}{22pt}\selectfont
\hrule
\halign{
\vrule#&\strut\quad\hfil#\hfil\quad&\vrule#&\quad\hfil#\hfil\quad
&\vrule#&\quad\hfil#\hfil\quad&\vrule#\cr
height 3pt&\omit&&\omit&&\omit&\cr
&Problem&&Possible&&Score&\cr\tline
height 3pt&\omit&&\omit&&\omit&\cr
&1&&10&&\emptybox&\cr\tline
&2&&16&&\emptybox&\cr\tline
&3&&20&&\emptybox&\cr\tline
&4&&15&&\emptybox&\cr\tline
&5&&12&&\emptybox&\cr\tline
&6&&12&&\emptybox&\cr\tline
&7&&15&&\emptybox&\cr\tline
&Extra Credit&&5&&\emptybox&\cr\tline
&Total&&100&&\emptybox&\cr
}\hrule}}}

\newpage
\begin{enumerate}
\item (10 points) Let $H$ and $K$ be subgroups of the group $G.$ Recall that $HK=\{hk \, : \, h \in H \text{ and } k \in K \}.$ Prove that {if $G$ is abelian}, then $HK \leq G.$\\
\vfill
\newpage

\item (16 points) Give an examples of the following, if they exist. Otherwise briefly explain why such examples do not exist.
	\begin{enumerate}
	\item a noncyclic group of order 13
	\vfill
	\item an infinite group $G$ and element $g \in G,$ such that $|g|$ is finite
	\vfill
	\item two nonisomorphic groups of order 18
	\vfill
	\item a nonabelian group $G$ and a subgroup $H$ of $G$ such that $H \lhd G$
	\vfill
	\end{enumerate}
\newpage

%%HW 5 #7
\item (20 points) 
	\begin{enumerate}
	\item Suppose that $H$ is a subgroup of the group $G$. Define the \emph{index} of $H$ in $G.$
	\vspace{1in}
	\item Give an example of a group $G$ and a \emph{nontrivial} subgroup $H$ such that the index of $H$ in $G$ is 50.
	\vspace{1in}
	\item Let $H$ be a subgroup of the group $G.$ Prove that if $[G : H]=2,$ then for every $a,b \in G\backslash H,$ $ab \in H.$\\	
	\end{enumerate}
\newpage
%%HW6 #9
\item (15 points)  Let $G$ be a group and $g \in G.$ Define the function $f_g(x) : G \to G$ by $f_g(x)=gxg^{-1}.$ Prove that $f_g$ is an isomorphism from $G$ to itself.

\newpage
%\item (10 point) Let $H$ be a subgroup of the group $G$ and let $a \in G.$ Prove that $aH \leq G$ if and only if $a \in H.$

\vfill

\item (12 points)
	\begin{enumerate}
	\item State Lagrange's Theorem
	\vfill
	\item Suppose $K$ is a proper subgroup of $H$ and $H$ is a proper subgroup of $G.$ (So $K<H<G.$) If $|K|=10$ and $|G|=200,$ what are the possible orders of $H$? (Or, what are the possible values for $|H|$?) Explain your reasoning.
	\vfill
	\end{enumerate}
\newpage
\item (12 points)
	\begin{enumerate}
	\item State the definition of a \emph{normal subgroup.}
	\vspace{1in}
	\item Let $G=GL_2(\bbR)$ and let $H=\{ A \in GL_2(\bbR) \,  : \, \text{det}(A)=2^k \text{ for } k \in \bbZ\}$ be a subgroup of $G.$ Prove that $H \lhd G.$ \\
	\emph{Note:} You do not need to prove that $H$ is a subgroup of $G.$ You only need to prove that $H$ is normal in $G.$
	\vfill
	\end{enumerate}

\newpage

\item (15 points) Let $G=\bbZ_{24}$ and $H=\langle 8 \rangle.$
	\begin{enumerate}
	\item In one or two sentences, explain why $H \lhd G.$
	\vfill
	\item List the distinct elements of $G/H.$
	\vfill
	\item For cosets $10+H$ and $20 +H$, determine $(10+H)+(20+H).$
	\vfill
	\item Determine the order of the element $10 + H$ in $G/H.$
	\vfill
	\item Can $G/H$ have an element of order 5? If so, find such an element. If not, explain why it is not possible.
	\vfill
	\end{enumerate}
\newpage
\end{enumerate}
\emph{Extra Credit:} (5 points) Suppose that $H$ and $K$ are subgroups of the group $G.$ Prove that if there exist elements $a,b \in G$ such that $aH \subseteq bK,$ then $H \subseteq K.$

\end{document}